### 2018-12-20 index.Rmd
update

### Toolbox for Automated Foehn Classification based on Mixture Models

The **foehnix** package provides a toolbox for automated probabilistic foehn
wind classification based on two-component mixture models (**foehn**
m**ix**ture models).  This method has first been proposed by [Plavcan et al.
(2015)](#plavcan2014) and compared to another semi-automatic classification,
foehn experts, students, and weather enthusiasts in the "Community Foehn
Classification Experiment" ([Mayr 2019](mayr2018)).

Foehn mixture models are a special case of the general flexible mixture model
class ([Fraley 2002](#fraley2000), [Leisch 2004](#leisch2004), [Gr&uuml;n
2007](#gruen2007), [Gr&uuml;n 2008](#gruen2008)), an unsupervised statistical
model to identify unobserveable clusters or components in data sets.
**foehnix** allows to estimate two-component mixture
models with additional concomitants.

Some of the features:

* Gaussian or logistic components with optional censoring or truncation.
* Concomitant variables for the probability model.
* Model assessment based on graphical output and information criteria.
* Automatic handling of missing values in the data set.

```{r, echo = FALSE, fig = TRUE, results = 'hide', fig.width = 12, fig.height = 8}
suppressPackageStartupMessages(library("foehnix"))
# Loading observations (data.frame), convert to zoo
data("ellboegen", package = "foehnix")
data("sattelberg", package = "foehnix")
timefun    <- function(x) as.POSIXct(x,  origin = "1970-01-01", tz = "UTC")
ellboegen  <- zoo(ellboegen[,-1],  timefun(ellboegen$timestamp))
sattelberg <- zoo(sattelberg[,-1], timefun(sattelberg$timestamp))
names(sattelberg) <- paste0("crest_", names(sattelberg))
data <- merge(ellboegen, sattelberg)
data$diff_t <- data$crest_t + 10.27 - data$t
windfilter <- list(dd = c(43, 223), crest_dd = c(90, 270))
mod <- foehnix(diff_t ~ ff, data = data, windfilter = windfilter, switch = TRUE)
# Plot
tsplot(mod, start = "2010-02-10", end = "2010-02-20")
```

### References

<p id="mayr2018">
Mayr GJ, Plavcan D, Laurence A, Elvidge A, Grisogono B, Horvath K, Jackson P, Neururer A, Seibert P, Steenburgh JW, Stiperski I, Sturman A, Ve&#269;enaj &#381;, Vergeiner J, Vosper S, Z&auml;ngl G (2018).
The Community Foehn Classification Experiment.
_Bulletin of the American Meteorological Society_, **99**(11), 2229&mdash;2235,
<a href="https://doi.org/10.1175/BAMS-D-17-0200.1" target="_blank">10.1175/BAMS-D-17-0200.1</a>
</p>

<p id="plavcan2014">
Plavcan D, Mayr GJ, Zeileis A (2014).
Automatic and Probabilistic Foehn Diagnosis with a Statistical Mixture Model.
_Journal of Applied Meteorology and Climatology_, **53**(3), 652&mdash;659,
<a href="https://dx.doi.org/10.1175/JAMC-D-13-0267.1" target="_blank">10.1175/JAMC-D-13-0267.1</a>
</p>

<p id="gruen2008">
Gr&uuml;n B, Friedrich L (2008).
FlexMix Version 2: Finite Mixtures with Concomitant Variables and Varying and Constant Parameters.
_Journal of Statistical Software, Articles_, **28**(4), 1&mdash;35,
doi:<a href="https://dx.doi.org/10.18637/jss.v028.i04" target="_blank">10.18637/jss.v028.i04</a>
</p>

<p id="gruen2007">
Gr\"un B, Leisch F (2007).
Fitting Finite Mixtures of Generalized Linear Regressions in _R_.
_Computational Statistics \& Data Analysis_, **51**(11),
doi:<a href="https://dx.doi.org/10.1016/j.csda.2006.08.014" target="_blank">10.1016/j.csda.2006.08.014</a>
</p>

<p id="leisch2004">
Friedrich L (2004).
FlexMix: A General Framework for Finite Mixture Models and Latent Class Regression in _R_.
_Journal of Statistical Software, Articles_, **11**(8), 1&mdash;18,
doi:<a href="https://dx.doi.org/10.18637/jss.v011.i08" target="_blank">10.18637/jss.v011.i08</a>
</p>

<p id="fraley2000">
Fraley C, Raftery AE (2000).
Model-Based Clustering, Discriminant Analysis, and Density Estimation.
_Journal of the American Statistical Association_, **97**(458), 611&mdash;631,
doi:<a href="https://dx.doi.org/10.1198/016214502760047131" target="_blank">10.1198/016214502760047131</a>
</p>